---
title: "Intro R"
author: "Witek ten Hove"
date: "`r format(Sys.time(), '%d-%m-%Y')`"
output:
  html_notebook: default
  word_document: default
---

## Story
Lara is a quality manager at a multinational company. One of her responsibilities is to monitor compliance to safety rules. She uses data that she receives every week from all the production faciclities. Local managers fill in a survey and Lara uses the aggregated answers to follow developments and see whether rules are implemented according to plans. Lara produces a monthly report which she sends to the head of Risk Management.

How can we help Lara to do her work more efficiently and generate better insights?

## Load libraries
We will be using some functions that are not part of the base R installation by loading the packages that contain them. For that we'll use the `library()` function.

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(gsheet)
library(tidyverse)
library(plotly)
```

## Read data

The data is gathered using Google Forms. We can read this data directly from the source.

```{r}
rawDF <- gsheet2tbl("https://docs.google.com/spreadsheets/d/1sZLHr2ebgieI9Qw3WDY4YFP0zKOMVSEbthdwjPobghE/edit?usp=sharing")[-1]
head(rawDF)
```

## Assess data quality
We can view the first few rows and get some basic statistical parameters (central tendency and spread).
```{r}
head(rawDF)
summary(rawDF)
```
The first column (variable) contains strings (type = `chr`). Most likely they are the IDs of the different facilities. We will change this type into categorical. Using the function `levels()` we can check all the different IDs. Apparently Lara added targets to the data.

We should also change the format of the columns with dates to type date`. Otherwise it will be handled as text data.

```{r}
rawDF$id <- as.factor(rawDF$id)
levels(rawDF$id)
rawDF$date <- as.Date(rawDF$date)
str(rawDF)
```

## Visualization
Now we can inspect the data visually. First we'll look at the data of one facility.

```{r}
flrDF <- filter(rawDF, id == "USA1")
plot(flrDF$date, flrDF$value)
```

Although informative the visualization is rather unattractive. The pacakge `ggplot2` (part of `tidyverse`) has functions for building nice visualizations. We will plot all data in one chart.

```{r}
p <- ggplot(rawDF) +
  geom_line(aes(x = date, y = value, group = id, color = id)) +
  ylim(0, 4) +
  scale_y_continuous(name = "level", breaks=c(1, 2, 3, 4), labels = c("BASIC", "DEVELOPMENT",
                                                                         "MATURITY", "EXCELLENCE"))
p
```

Using the package `plotly` we can add interaction (move the cursor over the chart).

```{r}
ggplotly(p)
```

